Bumble bee foraging bias

### 2021-01-11 analyses/visitation_bias.Rmd
syncing to office desktop, got rid of lots of old scripts and output

## And one for niche overlap
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

# 1. Is bumble bee foraging biased? How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models? Are some BB species more biased than others?
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
no_plant.sp <- niche_overlapper(visit_cover, plant.sp)

bias_plantsp.sitedate <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.sp", n()))) %>%
  left_join(no_plant.sp)

# Grouped by k.type.ss
no_ktype <- niche_overlapper(visit_cover, k.type.ss)

bias_ktype.sitedate <- biaseval(visit_cover, k.type.ss, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("ktype", n()))) %>%
  left_join(no_ktype)

# Grouped by plant.family
no_plant.family <- niche_overlapper(visit_cover, plant.family)

bias_family.sitedate <- biaseval(visit_cover, plant.family, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.family", n()))) %>%
  left_join(no_plant.family)

# Collated
bias_sitedate <- bind_rows(bias_plantsp.sitedate, 
                           bias_ktype.sitedate, 
                           bias_family.sitedate) %>%
  mutate(bb.sp = factor(bb.sp),
         yday = yday(date),
         elev = elev.mean/10) %>%
  left_join(no)

# When bias is high, it could mean that there are few good options on which all (or many) BB species converge. But apparently it doesn't. When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge. 
ggplot(bias_sitedate, aes(dprime, mean.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias_sitedate, aes(dprime, max.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias_sitedate, aes(log(total.flower.cover), dprime, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias_sitedate, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

brm.bias_niche <- brm(mean.niche.overlap ~ 1 + dprime + (1 | bb.sp),
                     cores = 3,
                     data = filter(bias_sitedate, criterion == "ktype"))

## 1.2 Visualize
```{r}
# Compare mean bias by criterion
ggplot(bias_sitedate, aes(criterion, dprime)) +
  geom_boxplot()

# Bias ~ elevation * criterion
ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot()

ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

# Does bias vary with abundance?
ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bias_sitedate, aes(bb.sp.abund, dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, )

ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth()

@@

## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
```{r}
# brms
brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 3,
                  iter = 2000,
                  file = "./output/brm_ktype1",
                  data = filter(bias_sitedate, criterion ==  "ktype"))

@@

brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "./output/brm_ktype2b",
                  data = filter(bias_sitedate, criterion ==  "ktype"))

@@

brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  family = "poisson",
                  file = "./output/brm_ktype3",
                  data = filter(bias_sitedate, criterion ==  "ktype"))

@@

lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
                  data = filter(bias_sitedate, criterion ==  "ktype"))
summary(lmer.ktype)
pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")

@@

# mgcv
gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
                   t2(yday, elev.mean, bs = c("tp", "gp")),
                 method = "fREML",
                 discrete = TRUE,
                 data = filter(bias_sitedate, criterion ==  "ktype")) %>% getViz()

@@

inla.ktype <- inla(dprime ~ elev.mean, 
                   family = "gaussian",
                   data = filter(bias_sitedate, criterion ==  "ktype"))

inla.pen <- inla(dprime ~ 1 + elev.mean,
  data = filter(bias_sitedate, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
summary(inla.pen)
```

## 1.6 Define hypotheses
```{r}
# Is d' different from 0?
H1 <- paste("Intercept = 0")

@@

## 1.7 Test contrasts
```{r}
### Plant.sp
hypothesis(brm.plantsp, H1)

@@

# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
## 2.1 Calculate bias per plant speces/ktype/family

@@

```{r}
# Function for calculating bias per group (= criterion)
biaseval_group <- function(x, criterion) {

@@

  x %>%
    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
    group_by(site, date, bb.sp, criterion) %>%
    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
              visits = sum(visits)) %>%
    group_by(site, date, bb.sp) %>%
    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
           prop.visits = visits/sum(visits)) %>%
    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
           bias.fold = bias/prop.flower.cover, # bias as fold change
           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
    mutate(yday = yday(date)) %>%
    select(site, date, yday, bb.sp, criterion, 
           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
           visits, flower.cover, prop.visits, prop.flower.cover)
}

bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
  left_join(site_data) %>%
  rename(group = criterion) %>%
  mutate(criterion = rep("plant.sp", n()))

@@

## 2.2 Visualize
```{r}
# Check the distribution of bias
ggplot(bias_group_ktype, aes(bias)) +
  geom_histogram()

@@

# Bias by Kugler morphotype
ggplot(bias_ktype, aes(criterion, bias)) +
  geom_boxplot()

ggplot(bias_ktype, aes(criterion, bias)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(bias_ktype, aes(bb.sp, bias)) +
  geom_boxplot() +
  facet_wrap(~criterion)

ggplot(bias_ktype, aes(elev.mean, bias)) +
  #geom_point() +
  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
  facet_wrap(~criterion)

# Bias by plant family
ggplot(bias_family, aes(criterion, bias)) +
  geom_boxplot()

ggplot(bias_family, aes(criterion, bias)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(bias_family, aes(elev.mean, bias)) +
  #geom_point() +
  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
  facet_wrap(~criterion)

ggplot(bias_family, aes(elev.mean, bias)) +
  #geom_point() +
  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
  facet_wrap(~criterion)

## 2.3 Explaining representation by floral traits, BB spp, and elevation (and maybe time)
```{r}
repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
                    data = bias_ktype)

@@

# 4. Can we explain BB-FL trait associations in terms of floral reward or handling time? We can use Neumayer and Paulus' (1999) data. 
## 4.1 
## 4.2 Do there appear to be suboptimalities in BB-FL trait associations? That is, are trait classes over/under-represented relative to their reward/efficiency? If so, how might explain such suboptimalities?  
  - Innate bias trumps learning?
  - Ecological modulators are involved, e.g. competition or distribution (see Sowig 1989)

>>>>

# 1. Is bumble bee foraging biased? How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models? Are some BB species more biased than others?
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.sp", n()))) %>%
  left_join(no_plant.sp)

# Grouped by k.type.ss
bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("ktype", n()))) %>%
  left_join(no_ktype)

# Grouped by plant.family
bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.family", n()))) %>%
  left_join(no_plant.family)

# Collated
bias <- bind_rows(bias_plantsp.sitedate, 
                  bias_ktype.sitedate, 
                  bias_family.sitedate) %>%
  mutate(bb.sp = factor(bb.sp),
         yday = yday(date),
         elev = elev.mean/10) 
```

## 1.2 Visualize
```{r}
# Compare mean bias by criterion
ggplot(bias, aes(criterion, dprime)) +
  geom_boxplot()

# Bias ~ elevation * criterion
ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
ggplot(bias, aes(year, dprime, color = criterion)) +
  geom_boxplot()

ggplot(bias, aes(year, dprime, color = criterion)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

# Does bias vary with abundance?
ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(bb.sp.abund, dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, )

ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth()

@@

## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
```{r}
# brms
brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 3,
                  iter = 2000,
                  file = "./output/brm_ktype1",
                  data = filter(bias, criterion ==  "ktype"))

@@

brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "./output/brm_ktype2b",
                  data = filter(bias, criterion ==  "ktype"))

@@

brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  family = "poisson",
                  file = "./output/brm_ktype3",
                  data = filter(bias, criterion ==  "ktype"))

@@

lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
                  data = filter(bias, criterion ==  "ktype"))
summary(lmer.ktype)
pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")

@@

# mgcv
gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
                   t2(yday, elev.mean, bs = c("tp", "gp")),
                 method = "fREML",
                 discrete = TRUE,
                 data = filter(bias, criterion ==  "ktype")) %>% getViz()

@@

inla.ktype <- inla(dprime ~ elev.mean, 
                   family = "gaussian",
                   data = filter(bias, criterion ==  "ktype"))

inla.pen <- inla(dprime ~ 1 + elev.mean,
  data = filter(bias, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
summary(inla.pen)
```

## 1.4 Define hypotheses
```{r}
# Is d' different from 0?
H1 <- paste("Intercept = 0")

@@

## 1.5 Test contrasts
```{r}
### Plant.sp
hypothesis(brm.plantsp, H1)

@@

# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented

## 2.1 Defining bias

@@

## 2.2 Function for calculating bias per group (= criterion)
```{r}
biaseval_group <- function(x, criterion) {

@@

  x %>%
    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
    group_by(site, date, bb.sp, criterion) %>%
    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
              visits = sum(visits)) %>%
    group_by(site, date, bb.sp) %>%
    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
           prop.visits = visits/sum(visits)) %>%
    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
           bias.fold = bias/prop.flower.cover, # bias as fold change
           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
    mutate(yday = yday(date)) %>%
    select(site, date, yday, bb.sp, criterion, 
           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
           visits, flower.cover, prop.visits, prop.flower.cover)
}
```

## 2.3 Calculate bias per plant speces/ktype/family
```{r}
bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
  left_join(site_data) %>%
  rename(group = criterion) %>%
  mutate(criterion = rep("plant.sp", n()))

@@

## 2.4 Visualize
```{r}
# Check the distribution of bias
ggplot(bias_group_ktype, aes(bias)) +
  geom_histogram()

@@

# Bias by Kugler morphotype
ggplot(bias_group_ktype, aes(group, bias)) +
  geom_boxplot()

ggplot(bias_group_ktype, aes(group, bias)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(bias_group_ktype, aes(bb.sp, bias)) +
  geom_boxplot() +
  facet_wrap(~group)

ggplot(bias_group_ktype, aes(elev.mean, bias)) +
  #geom_point() +
  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
  facet_wrap(~group)

# Bias by plant family
ggplot(bias_group_family, aes(group, bias)) +
  geom_boxplot()

ggplot(bias_group_family, aes(group, bias)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(bias_group_family, aes(bb.sp, bias)) +
  geom_boxplot() +
  facet_wrap(~group)

ggplot(bias_group_family, aes(elev.mean, bias)) +
  #geom_point() +
  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
  facet_wrap(~group)
```

## 2.5 Explaining bias by floral traits, BB spp, and elevation (and maybe time)
```{r}
repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
                    data = bias_ktype)

@@

# 4. How does bias relate to network structure?
## 4.1 A function for calculating niche overlap
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

## 4.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
```{r}
niche_overlap_plant.sp <- niche_overlapper(visit_cover, plant.sp)

niche_overlap_ktype <- niche_overlapper(visit_cover, k.type.ss)

niche_overlap_plant.family <- niche_overlapper(visit_cover, plant.family)

niche_overlap <- bind_rows(no_plant.sp, 
                           no_ktype, 
                           no_plant.family)

# When bias is high, it could mean that there are few good options on which all (or many) BB species converge. But apparently it doesn't. When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge. 
ggplot(bias, aes(dprime, mean.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias, aes(dprime, max.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias, aes(log(total.flower.cover), dprime, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

ggplot(bias, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

brm.bias_niche <- brm(mean.niche.overlap ~ 1 + dprime + (1 | bb.sp),
                     cores = 3,
                     data = filter(bias, criterion == "ktype"))

### 2021-01-11 writing/manuscript_v4.Rmd
syncing to office desktop, got rid of lots of old scripts and output

---
title: Bumble bee foraging bias
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter 
bibliography: refs.bib
output: html_document
---

@@

*mountain slopes give us floristic variation and abiotic variation while controlling for genetic variation within BB species, thus enabling us to study the lability of bias and partition it between innate and learned components*

The study bias in bumble bee foraging

Local elevation gradients along steep mountain slopes provide the opportunity to study broad environmental variation while controlling for potential confounding effects associated with geographic distance.

>>>>

---
title: "Bumble bee foraging bias"
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter
bibliography: refs.bib
output:
  pdf_document: default
  html_document: default
---

@@

The topic of floral preference in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables more confident causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions. Observational field studies, however, have tended to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il]. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic conditions while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underly floral biases and the network structures that result from them. Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude. In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make -->.

# Methods

## Field sampling

The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is characterized by mountain pastures mainly surrounded by coniferous forests. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.

Sampling consisted of repeated visits to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz.

## Data analysis

Prior to analysis, we aligned each bumble bee visitation sample with corresponding floral survey data. For most samples, floral surveying was conducted jointly with bumble bee visitation observation. In cases where bumble bee visitation was sampled without a joint floral survey, the bumble bee visitation data was paired with the nearest available floral survey date within a maximum distance of 7 days. Bumble bee visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis. In a few cases where more than one bumble bee visitation date was paired with a given floral survey date, the bumble bee visitation data were pooled.

To quantify cumulative bias per bumble bee species, we used the discrimination (d’) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation from the predictions of a null model in which the probability of visitation is purely a function of floral abundance.

# Results

# Discussion
